Constructing Sailing Match Race Schedules: Round Robin Pairing Lists

### 2015-03-24 paper.tex
Use showframe for now
Don't muck around with page numbers
Extra comma
Major layout tweaking

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathtools}
\usepackage[rightcaption]{sidecap}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{coffee2}

@@

\newcommand{\eachOccursExactlyOnce}{\mathrm{eachOccursExactlyOnce}}
\newcommand{\allDifferent}{\mathrm{allDifferent}}
\newcommand{\minimise}{\mathrm{minimise}}
\newcommand{\maximum}{\mathrm{maximum}}

@@

@@

\pagestyle{plain}
\begin{document}

@@

The schedules are produced in four stages. The first stage produces a schedule
that respects criteria 1, 4, 11 and 12, and minimises criteria 10 (boat
changes). The second stage constructs a schedule that respects criteria 1, 4,
11, 12 (again), has criteria 10 as a hard constraint generated from first
stage, and minimises criteria 6 (balance). This results in a schedule that
attempts to minimise boat changes \emph{and} balance. The third stage satisfies
criteria 9, and is a simple translation of the schedule produced in stage 2.
The final stage, orients skippers within matches to address criteria 2, 3, 5
and 8.

@@

The first thing we model is a skipper. A skipper $\sigma$ has a temporal view
of his schedule in the array of constrained integer variables $\timeSlot[0]$
to $\timeSlot[{t-1}]$ (equation \ref{eq3} below), and a state view in the array of
constraint integer variables $\state[0]$ to $\state[{f-1}]$ (equation \ref{eq4}
below), i.e.\ $\state[i]$ is the state of this skipper in flight $i$ and
corresponds to timeslots $\timeSlot[{i*m}]$ to $\timeSlot[{m(i+1)-1}]$.

\begin{flalign}
& f = \lceil n(n-1)/m  \rceil \label{eq1} \\
& t = n(n-1)/2 \label{eq2} \\
& \forall_{\tau \in [0..t-1]} ~ \timeSlot[{\tau}] \in \{-2..t-1\} \setminus \{\sigma\} \label{eq3} \\
& \forall_{i \in [0..f-1]} ~ \state[{i}] \in \{\FIRST,\MID,\LAST,\BYE,\END\} \label{eq4} \\
& \forall_{i \in [0..f]} ~ \state[{i}] = \FIRST \iff \timeSlot[{i \cdot m}] \geq 0  \label{eq5} \\
& \forall_{i \in [0..f]} ~ \state[{i}] = \MID \iff  \exists_{j \in [i.m+1 ..  i.m+m-2]} ~ \timeSlot[{j}] \geq 0  \label{eq6} \\
& \forall_{i \in [0..f]} ~ \state[{i}] = \LAST \iff \timeSlot[{(i+1).m - 1}] \geq 0  \label{eq7} \\
& \forall_{i \in [0..f]} ~ \state[{i}] = \BYE \iff  \forall_{j \in [i.m ..  m.(i+1)-1]} ~ \timeSlot[j] = -1  \label{eq8} \\
& \forall_{i \in [0..f]} ~ \state[{i}] = \END \iff  \forall_{j \in [i.m ..  m.(i+1)-1]} ~ \timeSlot[j] = -2  \label{eq9} \\
& \eachOccursExactlyOnce(\timeSlot, \{0, \ldots, n-1\} \setminus \{\sigma\}\})
\end{flalign}

In (\ref{eq3}) the array $\timeSlot$ is defined and in (\ref{eq4}) $\state$ is
defined. If variable $\timeSlot[{i}] = k$ and $k \geq 0$ then this skipper is
in a match with skipper $k$ at time $i$. Variable $\state[{i}]$ gives the state
of this skipper in flight $i$. The skipper can be in a match in the first time
slot in the flight (\ref{eq5}), or in the middle of the flight\footnote{\ldots
actually, not first and not last.} (\ref{eq6}), or in the last time slot of the
flight (\ref{eq7}). If all time slots in a flight are equal to -1 then the
skipper is in a bye (\ref{eq8}), and if all time slots in a flight are equal to
-2 then the skipper has finished all his matches (\ref{eq9}).

We must then ensure that each skipper $\sigma$ is in a match with all other
skippers $\{0, \ldots, n-1\} \setminus \{sigma\}$. This may be done by imposing
a global cardinality constraint \cite{globCard} on the array $\timeSlot$. The
$\state$ variables are then used to impose match race criteria 4 (if last in
flight $i$ then not first in flight $i+1$), criteria 11 (if last in flight $i$
then not in a bye in flight $i+1$) and criteria 12 (if flight $i$ is a bye then
flight $i$ is not first). In addition, we have the rule that if a skipper has
finished his matches in flight $i$ (i.e.\ $\END$) then he has finished his matches
in flight $i+1$ (and so on).  These criteria are imposed on the $\state$
variables via a deterministic finite automata (DFA) constraint \cite{Pesant04}.

The constraints above are shown pictorially in Figure~\ref{skipper5}, and
corresponds to skipper 5 in the Table~\ref{tab1} schedule. The $\state$ array
is presented as coloured boxes, and below that is the $\timeSlot$ array. The
arrows represent the channeling constraints (\ref{eq5}) to (\ref{eq9}). The
$\state$ array is coloured with green representing $\LAST$, yellow for $\MID$, blue
for $\FIRST$, red for $\LAST$ and (not shown) pink for $\BYE$. The schedule of
Table~\ref{tab1} is reproduced with the matches for skipper 5 emboldened. To
the right of the schedule is schema for the DFA. Coloured arcs represent a
transition from one state to another. The DFA is encoded as a set of transition
objects $\langle q_{i},\iota,q_{j} \rangle$, where a transition is made from
state $q_{i}$ to state $q_{j}$ when encountering input $\iota$. In addition we
specify the accepting states, which are all states except the start state
(white) and the bye state (pink).

@@

The decision variables are $\timeVar[{0}]$ to $\timeVar[{t-1}]$, i.e.\ for each
time slot we decide what match takes place. A symmetry break is made at the top
of search by forcing the first round to contain the matches $(0,1), (2,3), \ldots$,
i.e.\ $\forall_{i \in [0 \ldots m-1]} ~ \match[{2i,2i+1}] = i$. The schedule
presented has 6 boat changes: there are no boat changes in flight 0 (first
flight), flight 1 skipper 6 has a boat change, flight 2 skipper 3 has a boat
change, flight 3 it's skipper 4, flight 4 skipper 6, flight 5 skipper 0 and
last flight skipper 1.

>>>>

% \usepackage{showframe}

\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{url}
\usepackage{graphicx}
\usepackage{float}
\usepackage{listings}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathtools}
\usepackage[rightcaption]{sidecap}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{coffee2}

@@

\newcommand{\eachOccursExactlyOnce}{\mathrm{eachOccursExactlyOnce}}
\newcommand{\allDifferent}{\mathrm{allDifferent}}
\newcommand{\validStateSequence}{\mathrm{validStateSequence}}
\newcommand{\minimise}{\mathrm{minimise}}
\newcommand{\maximum}{\mathrm{maximum}}

@@

\newcommand\nldots{\,\hbox to 0.8em{.\hss.\hss.}\,}

@@

\begin{document}

@@

The schedules are produced in four stages. The first stage produces a schedule
that respects criteria 1, 4, 11 and 12, and minimises criteria 10 (boat
changes). The second stage constructs a schedule that respects criteria 1, 4,
11, 12 (again), has criteria 10 as a hard constraint generated from first
stage, and minimises criteria 6 (balance). This results in a schedule that
attempts to minimise boat changes \emph{and} balance. The third stage satisfies
criteria 9, and is a simple translation of the schedule produced in stage 2.
The final stage orients skippers within matches to address criteria 2, 3, 5
and 8.

@@

The first thing we model is a skipper. A skipper $\sigma$ has a temporal view
of his schedule in the array of constrained integer variables $\timeSlot[0]$ to
$\timeSlot[{t-1}]$ (\eqref{var:timeSlot} below), and a state view in the array
of constraint integer variables $\state[0]$ to $\state[{f-1}]$
(\eqref{var:state} below), i.e.\ $\state[i]$ is the state of this skipper in
flight $i$ and corresponds to timeslots $\timeSlot[{i*m}]$ to
$\timeSlot[{m(i+1)-1}]$.

\begin{figure}
\setlength{\mathindent}{0pt}\setlength{\abovedisplayskip}{0pt}\setlength{\belowdisplayskip}{0pt}
\begin{framed}
\begin{align}
    &f = \lceil n(n-1)/m  \rceil \label{var:f} \tag{V1} \\
    &t = n(n-1)/2 \label{var:t} \tag{V2} \\
    &\forall \tau \in \{0 \nldots t-1\}: \timeSlot[{\tau}] \in \{-2 \nldots t-1\} \setminus \{\sigma\} \label{var:timeSlot} \tag{V3} \\
    &\forall i \in \{0 \nldots f-1\}: \state[{i}] \in \{\FIRST,\MID,\LAST,\BYE,\END\} \label{var:state} \tag{V4} \\
    &\forall i \in \{0 \nldots f\}: \label{C1} \tag{C1} \\
    &\begin{alignedat}{3}
    &\hspace*{1em} \state[{i}] = \FIRST &\ &\Leftrightarrow&&\ \timeSlot[{i \cdot m}] \geq 0 \\
    &\hspace*{1em} \state[{i}] = \MID   &\ &\Leftrightarrow&&\ \exists j \in \{i\cdot m+1 \nldots i \cdot m+m-2\} : \timeSlot[{j}] \geq 0 \\
    &\hspace*{1em} \state[{i}] = \LAST  &\ &\Leftrightarrow&&\ \timeSlot[{(i+1)m - 1}] \geq 0 \\
    &\hspace*{1em} \state[{i}] = \BYE   &\ &\Leftrightarrow&&\ \forall j \in \{i \cdot m \nldots m \cdot (i+1)-1\} : \timeSlot[j] = -1 \\
    &\hspace*{1em} \state[{i}] = \END   &\ &\Leftrightarrow&&\ \forall j \in \{i \cdot m \nldots m \cdot (i+1)-1\} : \timeSlot[j] = -2
    \end{alignedat} \nonumber \\
&\eachOccursExactlyOnce(\timeSlot, \{0, \nldots, n-1\} \setminus \{\sigma\}) \label{C2} \tag{C2} \\
&\validStateSequence(\state) \label{C3} \tag{C3}
\end{align}
\end{framed}
\caption{A skipper-oriented model. A copy of these variables and constraints is
created for each skipper $\sigma$. The number of skippers is $n$, and $m$ is
the number of matches.}
\end{figure}

In \eqref{var:timeSlot} the array $\timeSlot$ is defined and in
\eqref{var:state}, $\state$ is defined. If variable $\timeSlot[{i}] = k$ and $k
\geq 0$ then this skipper is in a match with skipper~$k$ at time~$i$. Variable
$\state[{i}]$ gives the state of this skipper in flight~$i$. The cases in
\eqref{C1} say that a skipper can be in a match in the first time slot in the
flight, or in the middle of the flight\footnote{\ldots actually, not first and
not last.}, or in the last time slot of the flight. Alternatively, if all time
slots in a flight are equal to $-1$ then the skipper is in a bye, and if all time
slots in a flight are equal to $-2$ then the skipper has finished all his
matches.

We must then ensure that each skipper $\sigma$ is in a match with all other
skippers $\{0, \nldots, n-1\} \setminus \{\sigma\}$. This is \eqref{C2}, which
may be implemented by imposing a global cardinality constraint \cite{globCard}
on the array $\timeSlot$.

The $\state$ variables are then used to impose match race criteria 4 (if last
in flight $i$ then not first in flight $i+1$), criteria 11 (if last in flight
$i$ then not in a bye in flight $i+1$) and criteria 12 (if flight $i$ is a bye
then flight $i$ is not first).  These criteria are imposed by \eqref{C3}, which
is a deterministic finite automata (DFA) constraint \cite{Pesant04}.  This
constraint also ensures that if a skipper has finished his matches in flight
$i$ (i.e.\ $\END$) then he has finished his matches in flight $i+1$ (and so
on).

The constraints above are shown pictorially in Figure~\ref{skipper5}, and
corresponds to skipper 5 in the Table~\ref{tab1} schedule. The $\state$ array
is presented as coloured boxes, and below that is the $\timeSlot$ array. The
arrows represent the channeling constraints \eqref{C1}. The
$\state$ array is coloured with green representing $\LAST$, yellow for $\MID$, blue
for $\FIRST$, red for $\LAST$ and (not shown) pink for $\BYE$. The schedule of
Table~\ref{tab1} is reproduced with the matches for skipper 5 emboldened. To
the right of the schedule is schema for the DFA. Coloured arcs represent a
transition from one state to another. The DFA is encoded as a set of transition
objects $\langle q_{i},\iota,q_{j} \rangle$, where a transition is made from
state $q_{i}$ to state $q_{j}$ when encountering input $\iota$. In addition we
specify the accepting states, which are all states except the start state
(white) and the bye state (pink).

@@

The decision variables are $\timeVar[{0}]$ to $\timeVar[{t-1}]$, i.e.\ for each
time slot we decide what match takes place. A symmetry break is made at the top
of search by forcing the first round to contain the matches $(0,1), (2,3), \nldots$,
i.e.\ $\forall_{i \in [0 \nldots m-1]} ~ \match[{2i,2i+1}] = i$. The schedule
presented has 6 boat changes: there are no boat changes in flight 0 (first
flight), flight 1 skipper 6 has a boat change, flight 2 skipper 3 has a boat
change, flight 3 it's skipper 4, flight 4 skipper 6, flight 5 skipper 0 and
last flight skipper 1.
